Probabilistic Catalogs for Crowded Stellar Fields

### 2013-03-10 paper/ms.tex
punctuation

Telescopes don't make catalogs \citep{2011EAS....45..351H}, they
measure the intensity field.  Viewed through the lens of probabilistic
inference, the goals of astronomy are to take the information in the
telescope-generated records of the intensity field and use this information
to obtain quantities of astronomical interest with as little loss as
possible.  Insertion of a catalog-generation step in the inference
pipeline between the raw imaging and the final astrophysical analyses
is potentially lossy; the hard decisions of catalog making destroy
information, at least in principle.  Probability theory suggests that it may be less
lossy to pass forward not a catalog but a probabilisitic
description of all the catalogs that could be consistent with the
imaging---a posterior probability in (enormously large) space of possible catalogs.
Essentially, the creation of a catalog is an
attempt to answer the question,
``Given the image we have obtained, what objects are present in the field and
what are their properties?''.
This article represents an attempt at implementing this ambitous goal in the specific
situation where the only objects in the field are stars or other point sources.

Beyond these philosophical concerns, there are practical issues:
Standard methods for constructing catalogs can have difficulty in some
challenging situations. For example, when multiple sources overlap partially
or completely, it can be difficult to determine how many sources are present,
and how much flux belongs to each source. In principle, the uncertainty about
the existence and properties of the objects can be significant and should be
propagated into any inferences about the stellar population. A Bayesian approach
that obtains the posterior distribution over catalog space (rather than a single
catalog estimate) has the potential to overcome these problems by deblending
objects when it is possible, and clearly indicating the uncertainty remaining
when it is not possible.

>>>>

Telescopes don't make catalogs \citep{2011EAS....45..351H}, they
measure the intensity field.  Viewed through the lens of probabilistic
inference, the goals of astronomy are to take the information in the
telescope-generated records of the intensity field and use this information
to obtain quantities of astronomical interest with as little loss as
possible.  Insertion of a catalog-generation step in the inference
pipeline between the raw imaging and the final astrophysical analyses
is potentially lossy. The hard decisions of catalog making destroy
information, at least in principle.  Probability theory suggests that it may be less
lossy to pass forward not a catalog but a probabilisitic
description of all the catalogs that could be consistent with the
imaging---a posterior probability distribution in the (enormously large) space of possible catalogs.
Essentially, the creation of a catalog is an
attempt to answer the question,
``Given the image we have obtained, what objects are present in the field and
what are their properties?''.
This article represents an attempt at implementing this ambitous goal in the specific
situation where the only objects in the field are stars or other point sources.

Beyond these philosophical concerns, there are practical issues;
standard methods for constructing catalogs can have difficulty in some
challenging situations. For example, when multiple sources overlap partially
or completely, it can be difficult to determine how many sources are present,
and how much flux belongs to each source. In principle, the uncertainty about
the existence and properties of the objects can be significant and should be
propagated into any inferences about the stellar population. A Bayesian approach
that obtains the posterior distribution over catalog space (rather than a single
catalog estimate) has the potential to overcome these problems by deblending
objects when it is possible, and clearly indicating the uncertainty remaining
when it is not possible.
